Paper reviews

### 2015-04-15 Rmarkdown/biology_background.Rmd
Minor formatting

The nucleosomes are interconnected by _linker DNA_ to form the macromolecule called chromatin. The linker sequences are between $20--60$ nucleotides in length, while approximately 147 nucleotides are wrapped around each nucleosome and a further 20 nucleotides wrapped around each additional H1/H5 histone \citep{Annunziato:2008ue}.

@@

Non-CpG methylation is very common in other organisms, such as plants. For example, \citet{Lister:2008bh} found that in the widely-studied _Arabidopsis Thaliana_ that $45\%$ of cytosines that displayed statistically significant evidence of methylation are at CHG or CHH loci. However, they also found that the level of methylation at non-CpG loci is typically lower than that observed at CpG dinucleotides.

@@

Whole-genome bisulfite-sequencing remains an expensive assay, which limits its use in studies involving large numbers of samples. Furthermore, depending on the choice of several sequencing parameters, approximately $35--72\%$ of reads will not contain any CpGs (Table \ref{tab:reads_withou_cpgs}), which might be considered a gross waste of resources for some experiments.

\begin{table}[h]
\centering
\begin{tabular}{@{}rc@{}}
\toprule
Sample              & Reads with no CpG \\ \midrule
ADS-iPSC            & $35 \%$                \\
ADS-adipose         & $37 \%$                \\
ADS                 & $39 \%$                \\
H9\_Laurent         & $39 \%$                \\
FF-iPSC\_19.11      & $48 \%$                \\
FF-iPSC\_19.7       & $49 \%$                \\
H1+BMP4             & $49 \%$                \\
FF-iPSC\_6.9        & $50 \%$                \\
IMR90-iPSC          & $50 \%$                \\
H9                  & $51 \%$                \\
FF-iPSC\_19.11+BMP4 & $52 \%$                \\
FF                  & $53 \%$                \\
IMR90\_r2           & $57 \%$                \\
IMR90\_r1           & $58 \%$                \\
H1\_r2              & $59 \%$                \\
H1\_r1              & $62 \%$                \\
HSF1                & $72 \%$                \\ \bottomrule
\end{tabular}
\caption{The percentage of reads that do not contain any CpGs for each sample in the \emph{Lister} dataset. Paired-end reads are counted as a single read and as such have a lower percentage of reads without CpGs. See Chapter \ref{chap:datasets} for a description of the \emph{Lister} dataset.}
\label{tab:reads_withou_cpgs}
\end{table}

@@

Almost all bisulfite-based assays require a considerable amount of DNA, e.g., $1--5$ $\mu$g for whole-genome bisulfite-sequencing, which is extracted from a population of cells. However, the minimal amount of DNA is being reduced with each technological advance. Recently, both \citet{Guo:2013ih} and \citet{Smallwood:2014kn} published single-cell bisulfite-sequencing assays, although they are not yet a commercially available sequencing assay. Furthermore, these are not yet proper genome-wide assays. The technique of \citet{Guo:2013ih} is adapted from RRBS and so only assays a small percentage of cytosines in the genome. And the technique of \citet{Smallwood:2014kn}, while intended as a genome-wide assay, can reportedly only "accurately measure DNA methylation at up to 48.4% of cpG sites" \citep{Smallwood:2014kn}.

>>>>

The nucleosomes are interconnected by _linker DNA_ to form the macromolecule called chromatin. The linker sequences are between $20$ to $60$ nucleotides in length, while approximately 147 nucleotides are wrapped around each nucleosome and a further 20 nucleotides wrapped around each additional H1/H5 histone \citep{Annunziato:2008ue}.

@@

Non-CpG methylation is very common in other organisms, such as plants. For example, \citet{Lister:2008bh} found that in the widely-studied _Arabidopsis thaliana_ that $45\%$ of cytosines that displayed statistically significant evidence of methylation are at CHG or CHH loci. However, they also found that the level of methylation at non-CpG loci is typically lower than that observed at CpG dinucleotides.

@@

Whole-genome bisulfite-sequencing remains an expensive assay, which limits its use in studies involving large numbers of samples. Furthermore, depending on the choice of several sequencing parameters, approximately $35$ to $72\%$ of reads will not contain any CpGs (Table \ref{tab:reads_without_cpgs}), which might be considered a gross waste of resources for some experiments.

\begin{table}[h]
\centering
\begin{tabular}{@{}rc@{}}
\toprule
Sample              & Reads with no CpG \\ \midrule
ADS-iPSC            & $35 \%$                \\
ADS-adipose         & $37 \%$                \\
ADS                 & $39 \%$                \\
H9\_Laurent         & $39 \%$                \\
FF-iPSC\_19.11      & $48 \%$                \\
FF-iPSC\_19.7       & $49 \%$                \\
H1+BMP4             & $49 \%$                \\
FF-iPSC\_6.9        & $50 \%$                \\
IMR90-iPSC          & $50 \%$                \\
H9                  & $51 \%$                \\
FF-iPSC\_19.11+BMP4 & $52 \%$                \\
FF                  & $53 \%$                \\
IMR90\_r2           & $57 \%$                \\
IMR90\_r1           & $58 \%$                \\
H1\_r2              & $59 \%$                \\
H1\_r1              & $62 \%$                \\
HSF1                & $72 \%$                \\ \bottomrule
\end{tabular}
\caption{The percentage of reads that do not contain any CpGs for each sample in the \emph{Lister} dataset. Paired-end reads are counted as a single read and as such have a lower percentage of reads without CpGs. See Chapter \ref{chap:datasets} for a description of the \emph{Lister} dataset.}
\label{tab:reads_without_cpgs}
\end{table}

@@

Almost all bisulfite-based assays require a considerable amount of DNA, e.g., $1$ to $5$ $\mu$g for whole-genome bisulfite-sequencing, which is extracted from a population of cells. However, the minimal amount of DNA is being reduced with each technological advance. Recently, both \citet{Guo:2013ih} and \citet{Smallwood:2014kn} published single-cell bisulfite-sequencing assays, although they are not yet a commercially available sequencing assay. Furthermore, these are not yet proper genome-wide assays. The technique of \citet{Guo:2013ih} is adapted from RRBS and so only assays a small percentage of cytosines in the genome. And the technique of \citet{Smallwood:2014kn}, while intended as a genome-wide assay, can reportedly only "accurately measure DNA methylation at up to 48.4% of cpG sites" \citep{Smallwood:2014kn}.

### 2015-04-15 Rmarkdown/paper_review.Rmd
Pernickety.
Minor updates.
Minor formatting

## \cite{Cokus:2008fc}
\cite{Cokus:2008fc} published the BS-seq protocol for performing WGBS. Most of the results in this paper used BS-seq of wild-type _Arabidopsis Thaliana_ and DNMT mutants. They also report limited results from low-coverage BS-seq of mouse germ cells but these are not relevant to my discussion.

@@

\cite{Lister:2008bh} published the methylC-seq protocol for performing WGBS. Like \cite{Cokus:2008fc}, \cite{Lister:2008bh} studied wild-type and DNMT-defective mutant _Arabidopsis Thaliana_ strains. Also like \cite{Cokus:2008fc}, much of the analysis is by visual comparison of summarised data.

@@

\cite{Lacey:2013iy} details a modeling and inference framework for RRBS data, as well as a method for simulating RRBS data. Here I only describe the modelin framework and reserve the description and discussion of the simulation framework for __CHAPTER__.

@@

## \cite{Rackham:2015bv}

@@

>>>>

## \cite{Cokus:2008fc}
\cite{Cokus:2008fc} published the BS-seq protocol for performing WGBS. Most of the results in this paper used BS-seq of wild-type _Arabidopsis thaliana_ and DNMT mutants. They also report limited results from low-coverage BS-seq of mouse germ cells but these are not relevant to my discussion.

@@

\cite{Lister:2008bh} published the methylC-seq protocol for performing WGBS. Like \cite{Cokus:2008fc}, \cite{Lister:2008bh} studied wild-type and DNMT-defective mutant _Arabidopsis thaliana_ strains. Also like \cite{Cokus:2008fc}, much of the analysis is by visual comparison of summarised data.

@@

\cite{Lacey:2013iy} details a modeling and inference framework for RRBS data, as well as a method for simulating RRBS data. Here I only describe the modeling framework and reserve the description and discussion of the simulation framework for __CHAPTER__.

@@

## \citet{Rackham:2015bv}

@@

### \citet{Rackham:2015bv}

Simulate WGBS.
